---
title: "Eating Out"
author: "Ever De Leon"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "Spring 2024"    
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Who dines out more often?
There were two years in which the General Social Survey asked about the frequency of eating out. We look at one of them.

The variable name of interest is "eatout."  The URL is https://gss.norc.org/ In GSS data explorer look for "eatout." 

## Find those years 
For the first year the question was asked (1996) we have data. Your assignment is to analyze the GSS yearly data for 1996 and write a statistical report that answers the question.  The data are in the Files section.   

## introduction
$H_0$ There is no difference among the various races with regard to frequency of eating out  
$H_A$ There is a statistical difference between at least two races with regard to the frequency of eating out  


```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(haven)
require(knitr)
require(foreign)
```
```{r, eval=TRUE, warning=FALSE}
GSS <-read_dta(file="GSS1996.dta")
GSS <- readstata13::read.dta13(file="GSS1996.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,age,eatout)
mydata <- na.omit(mydata)
```

## Methods

Both 'race' and 'eatout' are categorical. Therefore we will use the Cat_Cat.rmd  helper file. 

## Descriptive Results

### Graphical Description

```{r}
barchartGC(~eatout+race, data=GSS)
barchartGC(~eatout+race, data=GSS, type="percent")
```


In this plot chart we can see that the overall frequency of the white race to eat out is higher than the black and the other races. The black race has the highest frequency for not eating out, compared to the white and other races. All the other races represented in the graph had a median statistics compared to the data from th white and black races.


## Numerical Description

```{r}
table1 <- xtabs(~race + eatout, data=mydata)
rowPerc(table1)
colPerc(table1)
```

The white race is eating at a percentage of 82.83%, the black race is eating at the percentage of 68.75%, and the other races are eating at the percentage of 77.78%. 

## Inferential Results


### Chi-squared Test Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

By looking at this data, we notice there is a statistical difference between the data so we will use a fisherxy test.  


### fisherXY Test Results

```{r}
fisherXY(mydata$race,mydata$eatout,printAll = TRUE)
```

When we first ran the fisher XY command we got a low p-value only between white and black. However, when we ran it the other time, after adding the print, all equal, true part, we got two different other values that were significantly higher than the first one.


## Summary and Conclusions

In conclusion the data suggest that there is a difference in the frequency, and the percentage of the routes between the black and the white races. The data indicated that there is a statistically significant difference in the data between the two races. The graphical and  numerical results, both indicate that the white race eats out at a higher frequency and percentage than the black and other races. The critical P-value showed that the frequency of eating out is the difference between the white and black races. Therefore, we reject the $H_0$ that There is no difference among the various races with regard to frequency of eating out