---
title: "Eating Out"
author: "Ever De Leon"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "Spring 2024"    
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Who dines out more often?
There were two years in which the General Social Survey asked about the frequency of eating out. We look at one of them.

The variable name of interest is "eatout."  The URL is https://gss.norc.org/ In GSS data explorer look for "eatout." 

## Find those years 
For the first year the question was asked (1996) we have data. Your assignment is to analyze the GSS yearly data for 1996 and write a statistical report that answers the question.  The data are in the Files section.   

## introduction
$H_0$ There is no difference among the various races with regard to frequency of eating out  
$H_A$ There is a statistical difference between at least two races with regard to the frequency of eating out  


```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
#require(haven)
GSS=read.csv('GSS1996.csv')
GSS1996$race = factor(GSS1996$race, level=c('black','white','other'))
GSS1996$eatout = factor(GSS1996$eatout, levels=c('yes','no'))
```

## Methods

Both 'race' and 'eatout' are categorical. Therefore we will use the Cat_Cat.rmd  helper file. 

## Descriptive Results

### Graphical Description

```{r}
barchartGC(~eatout+race, data=GSS1996)
barchartGC(~eatout+race, data=GSS1996, type="percent")
```


The X-Axis is eatout (1=yes, 2=no) labels on the columns are (1=black, 2=white, 3=other) 

White race does not eat out more than does, while Black race eats out more than does not.

## Numerical Description

```{r}
table= xtabs(~eatout+race, data=GSS1996)
print(table)
```
Black race eats out more than White race and about the same rate as other

## Infrential Results

### Chi-squared Test Results

```{r}
chisq.test(table)
chisqtestGC(table)
```
There seems to be no relationship in the chi-squared test.

### fisherXY Test Results

```{r}
fisher.test(table)
fisherXY(GSS1996$race,GSS1996$eatout, printAll=T)
```

There is a significant difference between groups 1 and 2, but no significant difference between groups 1 and 3, or between groups 2 and 3.

## Summary and Conclusions

the data suggest that there is a statistical difference between at least two races with regard to the frequency of eating out. Specifically, Black individuals tend to eat out more frequently compared to White individuals. However, there is no significant difference in the frequency of eating out between Black and Other racial groups, or between White and Other racial groups. Because of this we must reject the $H_0$